Identifiers for Iterative Forecasting

### 2021-11-16 paper/paper.Rmd

Checksums have many advantages compared to alternative identifiers for individual data objects: they are (1) secure, (2) sticky, (3) portable (4) rot resistant, (5) cheap, (6) faciliate caching downloads, (7) facilitate caching workflows,

@@

Worflows: The use of content-based identifiers helps in large workflows in serveral ways. As discussed above, it (1) guarantee that the workflow only runs with identical data each time and (2) accelerates performance by avoiding re-downloading of content each time a workflow is run.  This approach is also important to more complex workflows where we can easily avoid re-computing expensive operations when relevant parts of a data analysis are unchanged.  One example is a forecasting workflow using NEON data.  NEON filenames frequently change without change to the underlying data.  A content-based system can avoid re-running parts of an analysis that would be re-executed under a location-based protocol [@neonstore]

Limitations:
Dynamic data, databases.  Data stored in dynamic structures and extracted on the fly can be difficult to operate on in a content-based structure.  However, it is important to note that such formats are also not considered best practice for long term archiving, as changes in software can render such data inaccessible.  Nor are these approaches required to achieve scale.  GBIF archives one large CSV.  NEON or MODIS examples instead rely on thousands of individual files.  This simplifies data transfer.  
Modern, high-performance systems like parquet are designed explicitly to take advantage of distributed file storage rather than a single continguous database file.  Such module approaches facilitates provenance tracing based on content hash.

@@

- Many alternatives use base64-encoded strings.  Because base64 includes special characters such as `/`, these identifiers must aso accept percent-encoded strings (RFC6920 discusses this for `ni://`, in which percent encoding is optional on certain characters but not on others). This means that even within a given hash algorithm, we can have considerable variation in the literal string used as an identifier.  RFC6920 introduces other considerations as well, such as noting that padding character `=` should be removed.  Some implementations of the `ni://` spec fail to handle all of these cases, illustrating the challenges that come with the grreater complexity of base64-encoded identifiers.

>>>>

Checksums have many advantages compared to alternative identifiers for individual data objects: they are (1) secure, (2) sticky, (3) portable (4) rot resistant, (5) cheap, (6) facilitate caching downloads, (7) facilitate caching workflows,

@@

Worflows: The use of content-based identifiers helps in large workflows in several ways.
As discussed above, it (1) guarantee that the workflow only runs with identical data each time and (2) accelerates performance by avoiding re-downloading of content each time a workflow is run.
This approach is also important to more complex workflows where we can easily avoid re-computing expensive operations when relevant parts of a data analysis are unchanged.  One example is a forecasting workflow using NEON data.
NEON filenames frequently change without change to the underlying data.
A content-based system can avoid re-running parts of an analysis that would be re-executed under a location-based protocol [@neonstore]

Limitations:
Dynamic data, databases.
Data stored in dynamic structures and extracted on the fly can be difficult to operate on in a content-based structure.
However, it is important to note that such formats are also not considered best practice for long term archiving, as changes in software can render such data inaccessible.
Nor are these approaches required to achieve scale.  GBIF archives one large CSV.  NEON or MODIS examples instead rely on thousands of individual files. 
This simplifies data transfer.  
Modern, high-performance systems like parquet are designed explicitly to take advantage of distributed file storage rather than a single continguous database file.
Such modular approaches facilitates provenance tracing based on content hash.

@@

- Many alternatives use base64-encoded strings.
Because base64 includes special characters such as `/`, these identifiers must also accept percent-encoded strings (RFC6920 discusses this for `ni://`, in which percent encoding is optional on certain characters but not on others).
This means that even within a given hash algorithm, we can have considerable variation in the literal string used as an identifier.
RFC6920 introduces other considerations as well, such as noting that padding character `=` should be removed. 
Some implementations of the `ni://` spec fail to handle all of these cases, illustrating the challenges that come with the grreater complexity of base64-encoded identifiers.
